\doctitle

### 2023-06-08 introduction.tex
Remove todos
[FINAL SUBMITTED] The very final version of my thesis. Compiles clean.

\edit{Add citations all over the intro}

Machine learning has emerged as an important tool in applied mathematics problems as data-driven approaches have begun to outperform  first-principle modeling in the tasks of classification, recognition, autonomous control, protein folding, and even video games. However, due to the absence of prior knowledge,  purely data-driven approaches typically require a large amount of data to approximate the behavior of first-principle models even in the simplest scenarios. Thus, the most promising modeling approaches of today are those that integrate both data-driven and first-principle components. These latter components frequently stem from domain expertise and pre-existing beliefs. Incorporating prior knowledge into a data-driven method is a formidable task that demands a considerable amount of ingenuity. This thesis serves to illustrate two distinct examples of such endeavors.  Together, these examples represent a diverse range of challenges in incorporating prior knowledge into a model and describe the strategies employed to surmount them.

An instance of such a problem is feature selection for mixed-effect models. Feature selection problems typically arise when one believes that a small subset of features within a dataset captures the majority of the variance of the quantity of interest; this subset of features is commonly known as sparse support. The canonical feature selection methods typically provide control over the  sparsity of the solution and output a candidate support for it. In the absence of additional information about the support apart from its probable sparsity, these methods furnish a satisfactory toolkit for practitioners.

@@

The second example serves to demonstrate how one can incorporate a degree of partial knowledge of physics into a deep neural network. Methods for data-driven modeling of physical phenomena have been increasingly popular in large part due to the advancements in GPU computations and automatic differentiation tools. These methods have proven particularly useful in reduced-order modeling tasks.  Using data, a neural network can find a representation of a system on a reduced-order manifold via a non-linear transformation from an observable space to that manifold. Classic techniques, such as Galerkin Proper Orthogonal Decomposition (POD) or Dynamic Mode Decomposition (DMD), are fundamentally incapable of finding such spaces because these methods are confined to linear, albeit optimal, transformations. However, the exceptional approximation capabilities of deep networks come at the cost of increased instability. In particular, the networks tend to identify manifolds that overfit the data and, consequently, fail at extrapolation tasks. Such deficiencies compromise the efficacy of the entire system in which the network is utilized, including applications in control, model identification, and compressive sensing.

Recent works have shown that one can improve a neural-network based model by incorporating partial knowledge of physics. For example, a model can find a better reduced-order manifold if it knows that it predicts the behavior of a shock wave that obeys Burger's equation. Chapter~\ref{ch:pinode} is devoted to the development of a framework for training Reduced-Order Models (ROMs) with Physics-Informed Neural Ordinary Differential Equations (PINODE). It illustrates how one can transfer knowledge from a known equation to a model that predicts solutions of said equation using collocation points.  The resulting ROMs are able to extrapolate forward in time considerably more effectively,  perform better for unseen initial conditions, and exhibit less sensitivity to noise. Finally, in Chapter~\ref{ch:cs} I demonstrate the utility of PINODE models in compressive sensing applications, where they effectively "stitch" together separate timeframes and go below the Nyquist sampling limit.

\paragraph{Attribution} Throughout my PhD I was fortunate to work with incredibly talented advisors and collaborators who significantly contributed to the content of this work. Chapter~\ref{ch:msr3} captures a sequence of joint works that were completed between 2019 and 2023 at the University of Washington's Institute for Health Metrics and Evaluations (IHME) together with Dr. Aleksandr Aravkin, Dr. James Burke, and Dr. Peng Zheng. The MSR3 framework algorithm was introduced in \cite{sholokhov2022relaxation}, with the theoretical foundations developed in \cite{aravkin2022jimtheory}, and the software implementation published in \cite{sholokhov2023pysr3}. Chapter~\ref{ch:pinode} represents my work at Mitsubishi Electric Research Labs (MERL) under the guidance of Dr. Hassan Mansour and Dr. Saleh Nabi. The PINODE framework was developed in \cite{sholokhov2023pinode} and is based on our earlier joint work with Dr. Yuying Liu on Physics-Informed Koopman Networks~(\cite{liu2022physics}). In particular, Figure~\ref{fig:pikn_burgers_compression} and the related example come from that work. Chapter~\ref{ch:cs} was developed in~\cite{sholokhov2023cs} together with Dr. Hassan Mansour, Dr. Saleh Nabi, Dr. J. Nathan Kutz, and Dr. Steven Brunton at the University of Washington. Many parts of the works listed above were taken verbatim to this document per the explicit permission of my co-authors.

>>>>

Machine learning has emerged as an important tool in applied mathematics problems as data-driven approaches have begun to outperform  first-principle modeling in the tasks of classification (\cite{lecun1999object}), recognition (\cite{deng2009imagenet,shafiq2022deep}),  autonomous control (\cite{janai2020computer}), protein folding (\cite{jumper2021highly}), and even video games (\cite{chen2016evolution}). However, due to the absence of prior knowledge,  purely data-driven approaches typically require a large amount of data to approximate the behavior of first-principle models (\cite{kutz2013data,pan2018long}) . Thus, the most promising modeling approaches of today are those that integrate both data-driven and first-principle components (\cite{rackauckas2020udes,raissi2019physics}). These latter components frequently stem from domain expertise and pre-existing beliefs. Incorporating prior knowledge into a data-driven method is a formidable task that demands a considerable amount of ingenuity. This thesis serves to illustrate two distinct examples of such endeavors.  Together, these examples represent a diverse range of challenges in incorporating prior knowledge into a model and describe the strategies employed to surmount them.

An instance of such a problem is feature selection for mixed-effect models. Feature selection problems typically arise when one believes that a small subset of features within a dataset captures the majority of the variance of the quantity of interest; this subset of features is commonly known as sparse support. The canonical feature selection methods, e.g. \cite{ibrahim2011fixed,Fan2012}, typically provide control over the  sparsity of the solution and output a candidate support for it. In the absence of additional information about the support apart from its probable sparsity, these methods furnish a satisfactory toolkit for practitioners.

@@

The second example serves to demonstrate how one can incorporate a degree of partial knowledge of physics into a deep neural network. Methods for data-driven modeling of physical phenomena have been increasingly popular in large part due to the advancements in GPU computations and automatic differentiation tools, such as pytorch (\cite{pytorch}), jax (\cite{jax2018github}), and Neural ODE frameworks (\cite{chen2018neuralode}). These methods have proven particularly useful in reduced-order modeling tasks (\cite{rackauckas2020udes}).  Using data, a neural network can find a representation of a system on a reduced-order manifold via a non-linear transformation from an observable space to that manifold. Classic techniques, such as Galerkin Proper Orthogonal Decomposition (POD) (\cite{fukunaga2013introduction}) or Dynamic Mode Decomposition (DMD) (\cite{schmid2022dynamic}), are fundamentally incapable of finding such spaces because these methods are confined to linear, albeit optimal, transformations (\cite{benner2015survey}). However, the exceptional approximation capabilities of deep networks come at the cost of increased instability. In particular, the networks tend to identify manifolds that overfit the data and, consequently, fail at extrapolation tasks. Such deficiencies compromise the efficacy of the entire system in which the network is utilized, including applications in control, model identification, and single-pixel imaging.

Recent works, e.g. \cite{raissi2019physics,liu2022physics}, have shown that one can improve a neural-network based model by incorporating partial knowledge of physics. For example, a model can find a better reduced-order manifold if it knows that it predicts the behavior of a shock wave that obeys Burger's equation. Chapter~\ref{ch:pinode} is devoted to the development of a framework for training Reduced-Order Models (ROMs) with Physics-Informed Neural Ordinary Differential Equations (PINODE). It illustrates how one can transfer knowledge from a known equation to a model that predicts solutions of said equation using collocation points.  The resulting ROMs are able to extrapolate forward in time considerably more effectively,  perform better for unseen initial conditions, and exhibit less sensitivity to noise. Finally, in Chapter~\ref{ch:cs} I demonstrate the utility of PINODE models in applications to single-pixel imaging (SPI), where they effectively "stitch" together separate timeframes and go below the Nyquist sampling limit.

\paragraph{Attribution} Throughout my PhD I was fortunate to work with incredibly talented advisors and collaborators who significantly contributed to the content of this work. Chapter~\ref{ch:msr3} captures a sequence of joint works that were completed between 2019 and 2023 at the University of Washington's Institute for Health Metrics and Evaluations (IHME) together with Dr. Aleksandr Aravkin, Dr. James Burke, and Dr. Peng Zheng. The MSR3 framework algorithm was introduced in \cite{sholokhov2022relaxation}, with the theoretical foundations developed in \cite{aravkin2022jimtheory}, and the software implementation published in \cite{sholokhov2023pysr3}. Chapter~\ref{ch:pinode} represents my work at Mitsubishi Electric Research Labs (MERL) under the guidance of Dr. Hassan Mansour and Dr. Saleh Nabi. The PINODE framework was developed in \cite{sholokhov2023pinode} (under peer review at the moment of writing this document) and is based on our earlier joint work with Dr. Yuying Liu on Physics-Informed Koopman Networks~(\cite{liu2022physics}). In particular, Figure~\ref{fig:pikn_burgers_compression} and the related example come from that work. Chapter~\ref{ch:cs} was developed in~\cite{sholokhov2023cs} together with Dr. Hassan Mansour, Dr. Saleh Nabi, Dr. J. Nathan Kutz, and Dr. Steven Brunton at the University of Washington, and is an ongoing work to date. Many parts of the works listed above were taken verbatim to this document per the explicit permission of my co-authors.

### 2023-06-08 msr3-theory.tex
[FINAL SUBMITTED] The very final version of my thesis. Compiles clean.

\begin{lemma}[$\LL+\phimu$ is Weakly Convex]\label{lem:LL weak cvx}
Let $\LL$ be as given in \eqref{eq:lme_diagonal_setup}. Then
\begin{equation}\label{eq:hess LL}
\nabla^2\LL{(\beta,\gam)}=\sum_{i=1}^m
S_i^T\begin{bmatrix}X_i^T\\ -Z_i^T\end{bmatrix}
\Omega_i(\gam)^{-1}
\begin{bmatrix}X_i& -Z_i\end{bmatrix}S_i
-\begin{bmatrix}
0&0\\ 0& \half(Z_i^T\Omega_i(\gam)^{-1}Z_i)^{\circ2}
\end{bmatrix},
\end{equation}
for all
$(\beta,\gam)\in\R^p\times\R^q_+$, where 
\[
S_i:=\begin{bmatrix}
I_q&0\\ 0&\Diag{Z_i^T\Omega_i^{-1}(X_i\beta-Y_i)}
\end{bmatrix}
\]
and, for any $A\in\R^{t\times t}$, $A^{\circ 2}:=A\circ A$. In particular, this implies
that the matrix
\begin{equation}\label{eq:psd for LL}
\begin{bmatrix}
\nabla_{\beta\beta}\LL(\beta,\gam)&\nabla_{\gam\beta}\LL(\beta,\gam)\\
\nabla_{\beta\gam}\LL(\beta,\gam)&\nabla_{\gam\gam}\LL(\beta,\gam)
+\blam I\end{bmatrix}
\end{equation}
is positive semidefinite
for $\bar\eta= \nu m$, where
\[
\nu:=max\lset{(1/2) \mu_\mmin(\Lam_i)^{-2}\sig^4_\mmax(Z_i)}{i=1,\dots,m},
\]
$\mu_\mmin(\Lam_i)$ is the smallest eigenvalue of $\Lam_i$, and
 $\sig_\mmax(Z_i)$ is the largest singular value of $Z_i,\, i=1,\dots,m$.
Consequently, for any $\tbg\in\dom{\tR}$ and $\mu\ge 0$, the mapping 
\(
\bg\mapsto \LL_{\emu}(\bg, \tbg)
%\hLL(\bg,\tgam) %:=\LL(\beta,\gam)+\frac{\blam}{2}\norm{\gam-\hgam}^2
%\sum_{i=1}^m\frac{\nu_i}{2}\norm{\gam-\hat\gam}^2,
\)
is convex for all $\eta\ge\bar\eta:=\nu m$.
In particular, this implies that $\LL+\phimu$ is weakly convex for any $\mu\ge 0$, 
and the mapping
\(
%\begin{aligned}
\bg\mapsto\LL_{\emu}(\bg,\tbg)
%&=
%\LL\bg+\phimu(\gam)+
%\keta(\beta-\tbeta,\beta-\tbeta)
%\\ &=
%\hLL(\bg,\tgam)+    \frac{\eta}{2}\norm{\bg-\tbg}^2 
%\end{aligned}
\)
is strongly convex for $\eta> \bar\eta$ with modulus of strong convexity $(\eta-\bar\eta)$
regardless of the choice of $\tbg\in\Rp\times\Rq$.
%Moreover, the mapping 
%\eq{\label{eq:defn L sub eta}
%(\beta, \gamma)\mapsto    
%\LL_{\eta}(x, w) := \tLL(x, w) + \frac{\eta}{2}\norm{x-w}^2 
%}
%is strongly convex for every $\eta>0$ and any choice of 
%$(\tbeta,\tgam)\in\R^p\times\R^q$.
\end{lemma}
\begin{proof}
The formula for $\nabla^2\LL$ is given in Appendix \ref{appendix:derivatives_of_lmm}
(see \eqref{eq:hess LL}).
By \cite[Theorem 3.1]{ABBP2021}, 
$\mu_\mmax(\left(Z_i^T\Omega_i(\gam)^{-1}Z\right)\le
\lam_\mmin^{-1}\sig_\mmax^2(Z_i)$, and since 
$\mu_\mmax(H^{\circ 2})\le \mu_\mmax^2(H)$
for all $H\in\bS^q_+$ \cite{HJ85}, we have
\[
\mu_\mmax\left(\half(Z_i^T\Omega_i(\gam)^{-1}Z_i)^{\circ2}\right)
\le (1/2) \lam_\mmin^{-2}
\sig^4_\mmax(Z_i)=:\nu_i
\qquad  i=1,\dots,m.
\]
This establishes that the matrix in \eqref{eq:psd for LL} is positive semidefinite.
Since $\phimu$ is convex, the mapping
\[
(\beta,\gam)\mapsto \LL_{\emu}(\bg,\tbg) %\hLL((\beta,\gam),\tgam)
\]
is strongly convex for any choice
of $\blam> \nu m$, where 
\eq{\label{eq:nu}
\nu:=\max_{i=1,\dots,m}\nu_i.
}
%The strong convexity statement follows from 
%the fact that
%$\nabla^2 \LL_{\eta}((\beta, \gamma), (\tbeta, \tgamma))
%=\nabla^2 \LL(\beta,\gam)+\eta I$ where $\nabla^2 \tLL(\beta,\gam)$
%is positive semidfinite.
\end{proof}
For the remainder of the paper, we assume that 
\eq{\label{eq:blam1}\eta> \nu m=:\bar\eta} 
so that the mapping
$(\beta,\gam)\mapsto \LL_{\emu}(\bg,\tbg)$ %\hLL((\beta,\gam),\tgam)$
is strongly convex with 
%$\nabla_{\bg\bg}^2\hLL(\bg,\tgam)$ 
positive definite Hessian
regardless of the 
choice of $\tgam\in\R^q$. With this in mind, the function $\uemu$ defined by
\eqref{eq:value_function_definition} resembles a Moreau envelope.
%of the convex function $\hLL+\phi_\mu$.
However, this is misleading since, in particular,
we are not even assured of the existence of solutions to
the optimization problem defining $\uemu$.

@@

\begin{theorem}[Convergence of Algorithms \ref{alg:MSR3} and \ref{alg:pgd with bt}]\label{thm:convergence}
Let the assumptions of Theorem \ref{thm:relaxed existence} hold, and
let $\Phi_\emu$ be as defined in \eqref{eq:varphi}. 
%and suppose
%$\gammax\in\R^q_{++}$. 
Let $\{\tbgk\}$ be a sequence generated
by either Algorithm \ref{alg:MSR3} or \ref{alg:pgd with bt} 
with parameters 
$\theta\in(0,1),\ \tau\in(0,1)$,
$\eta>0,\ \mu>0$, $\eps_{\mbox{\tiny Tol}}=0$,
$t_0>0$, and $\gammax>\tgam^0$. 
Then, given $\zeta>\uemu(\tbeta^0,\tgam^0)$ there is an 
$L_{(\emu,\gammax,\zeta)}>0$ such that $\nabla\uemu$ is 
$L_{(\emu,\gammax,\zeta)}$-smooth over 
$\widetilde\VV((\emu,\gammax,\zeta,1)$. 
In Algorithm \ref{alg:pgd with bt}, replace 
$L_\emu$ with $L_{(\emu,\gammax,\zeta)}$ and set
 \[
 \begin{aligned}
 M&:=\begin{cases}
 \alf(1-\alf\frac{L_{(\emu,\gammax,\zeta)}}{2}),&
 \mbox{in Algorithm \ref{alg:pgd with bt}},
 \\
 \frac{2t_0\theta^2(1-\tau)}{\max\{2\theta(1-\tau),t_0L_{(\emu,\gammax,\zeta)}\}},&
 \mbox{in Algorithm \ref{alg:MSR3}},
 \end{cases}
 \qquad\text{and}
 \\
 r&:=
 \begin{cases}
 \alf,&
 \mbox{in Algorithm \ref{alg:pgd with bt}},
 \\
t_0,&
 \mbox{in Algorithm \ref{alg:MSR3}}.
 \end{cases}
 \end{aligned}\]
Then either $\tgam^k>\gammax$ after a finite number of iterations and the algorithms terminate, or the 
following hold:
\begin{enumerate}
\item
The sequence $\Phi_\emu\tbgk$ is nondecreasing.
In addition, \[\Phi_\emu(\tbgk)\tbeta^{k+1},\tgam^{k+1})<\Phi_\emu\tbgk\]
if and only if $\tbgk$ is not a stationary point of \eqref{eq:relax3}.
\item
$\norm{\tbgk-\prox_{r\tR}(\tbgk-r\nabla u_\emu\tbgk)}\rightarrow 0$ with
\[
\min_{i=0,1,\dots,k}\norm{\tbgk-\prox_{r\tR}(\tbgk-r\nabla u_\emu\tbgk)}
\le \frac{\sqrt{\Phi_\emu(\tbeta^0,\tgam^0)-\Phi_\emu^\mathsf{OPT}}}
{\sqrt{M(k+1)}},
\]
where $\Phi_\emu^\mathsf{OPT}:=\inf \Phi_\emu$.
\item
All limit points of the sequence $\{\tbgk\}$ are stationary points of
problem \eqref{eq:relax3}.
\end{enumerate}
\end{theorem}
\begin{proof}
As observed prior to the statement of the theorem, 
both Algorithm \ref{alg:MSR3} and \ref{alg:pgd with bt}
behave as if they were applied to the the functions 
$\hat u_\emu$ and $\hR$ defined in \eqref{eq:hats}. 
It was also shown that the functions $\hat u_\emu$ and $\hR$
satisfy the Assumptions (A)-(C) required by \cite[Theorem 10.15]{AB17}.
Hence, the consequences of \cite[Theorem 10.15]{AB17} hold.
By translating the notion of \cite[Theorem 10.15]{AB17} to that of this paper,
we obtain the result.
\end{proof}

@@

\noindent
which is motivated by the expression for $\nabla^2\LL_{\emu}\bg$ given in
\eqref{eq:hess LL}. That is, we simply drop the negative semi-definite term
$-\sum_{i=1}^m\begin{bmatrix}
0&0\\ 0& \half(Z_i^T\Omega_i(\gam)^{-1}Z_i)^{\circ2}
\end{bmatrix}$.
With this modification, the subproblems we solve are strongly convex for all $\eta>0$.
Consequently, the problem of initializing $\eta$ is less problematic. Our numerical experiments indicate that the performance of the algorithm is robust with respect to 
$\eta$. For this reason, we choose an initial value for $\eta$ and 
then leave it fixed over all iterations. Our method for choosing $\eta$ is described
in \cite[Section 4, Figure 5]{Practice}. Briefly, we maximize the 
Baysian Information Criterion (BIC) over a grid of values for $\eta$. 
 The resulting BIC response curve shows 
that the method is robust with respect to the 
 choice of $\eta$ and choosing $\eta\in [1,10]$ yields 
 accurate solutions for our selected test problems. 
 Once $\eta$ is fixed the
 PGD algorithm can be applied to solve the problem \eqref{eq:relax3}
 for decreasing values of $\mu$.

>>>>

\begin{lemma}[$\LL+\phimu$ is Weakly Convex]\label{lem:LL weak cvx}
Let $\LL$ be as given in \eqref{eq:lme_diagonal_setup}. Then
\begin{equation}\label{eq:hess LL}
\nabla^2\LL{(\beta,\gam)}=\sum_{i=1}^m
S_i^T\begin{bmatrix}X_i^T\\ -Z_i^T\end{bmatrix}
\Omega_i(\gam)^{-1}
\begin{bmatrix}X_i& -Z_i\end{bmatrix}S_i
-\begin{bmatrix}
0&0\\ 0& \half(Z_i^T\Omega_i(\gam)^{-1}Z_i)^{\circ2}
\end{bmatrix},
\end{equation}
for all
$(\beta,\gam)\in\R^p\times\R^q_+$, where 
\[
S_i:=\begin{bmatrix}
I_q&0\\ 0&\Diag{Z_i^T\Omega_i^{-1}(X_i\beta-Y_i)}
\end{bmatrix}
\]
and, for any $A\in\R^{t\times t}$, $A^{\circ 2}:=A\circ A$. In particular, this implies
that the matrix
\begin{equation}\label{eq:psd for LL}
\begin{bmatrix}
\nabla_{\beta\beta}\LL(\beta,\gam)&\nabla_{\gam\beta}\LL(\beta,\gam)\\
\nabla_{\beta\gam}\LL(\beta,\gam)&\nabla_{\gam\gam}\LL(\beta,\gam)
+\blam I\end{bmatrix}
\end{equation}
is positive semidefinite
for $\bar\eta= \nu m$, where
\[
\nu:=max\lset{(1/2) \mu_\mmin(\Lam_i)^{-2}\sig^4_\mmax(Z_i)}{i=1,\dots,m},
\]
$\mu_\mmin(\Lam_i)$ is the smallest eigenvalue of $\Lam_i$, and
 $\sig_\mmax(Z_i)$ is the largest singular value of $Z_i,\, i=1,\dots,m$.
Consequently, for any $\tbg\in\dom{\tR}$ and $\mu\ge 0$, the mapping 
\(
\bg\mapsto \LL_{\emu}(\bg, \tbg)
%\hLL(\bg,\tgam) %:=\LL(\beta,\gam)+\frac{\blam}{2}\norm{\gam-\hgam}^2
%\sum_{i=1}^m\frac{\nu_i}{2}\norm{\gam-\hat\gam}^2,
\)
is convex for all $\eta\ge\bar\eta:=\nu m$.
In particular, this implies that $\LL+\phimu$ is weakly convex for any $\mu\ge 0$, 
and the mapping
\(
%\begin{aligned}
\bg\mapsto\LL_{\emu}(\bg,\tbg)
%&=
%\LL\bg+\phimu(\gam)+
%\keta(\beta-\tbeta,\beta-\tbeta)
%\\ &=
%\hLL(\bg,\tgam)+    \frac{\eta}{2}\norm{\bg-\tbg}^2 
%\end{aligned}
\)
is strongly convex for $\eta> \bar\eta$ with modulus of strong convexity $(\eta-\bar\eta)$
regardless of the choice of $\tbg\in\Rp\times\Rq$.
%Moreover, the mapping 
%\eq{\label{eq:defn L sub eta}
%(\beta, \gamma)\mapsto    
%\LL_{\eta}(x, w) := \tLL(x, w) + \frac{\eta}{2}\norm{x-w}^2 
%}
%is strongly convex for every $\eta>0$ and any choice of 
%$(\tbeta,\tgam)\in\R^p\times\R^q$.
\end{lemma}
\begin{proof}
The formula for $\nabla^2\LL$ is given in Appendix \ref{appendix:derivatives_of_lmm}
(see \eqref{eq:hess LL}).
By \cite[Theorem 3.1]{ABBP2021}, 
$\mu_\mmax(\left(Z_i^T\Omega_i(\gam)^{-1}Z\right)\le
\lam_\mmin^{-1}\sig_\mmax^2(Z_i)$, and since 
$\mu_\mmax(H^{\circ 2})\le \mu_\mmax^2(H)$
for all $H\in\bS^q_+$ \cite{HJ85}, we have
\[
\mu_\mmax\left(\half(Z_i^T\Omega_i(\gam)^{-1}Z_i)^{\circ2}\right)
\le (1/2) \lam_\mmin^{-2}
\sig^4_\mmax(Z_i)=:\nu_i
\qquad  i=1,\dots,m.
\]
This establishes that the matrix in \eqref{eq:psd for LL} is positive semidefinite.
Since $\phimu$ is convex, the mapping
\[
(\beta,\gam)\mapsto \LL_{\emu}(\bg,\tbg) %\hLL((\beta,\gam),\tgam)
\]
is strongly convex for any choice
of $\blam> \nu m$, where 
\eq{\label{eq:nu}
\nu:=\max_{i=1,\dots,m}\nu_i.
}
%The strong convexity statement follows from 
%the fact that
%$\nabla^2 \LL_{\eta}((\beta, \gamma), (\tbeta, \tgamma))
%=\nabla^2 \LL(\beta,\gam)+\eta I$ where $\nabla^2 \tLL(\beta,\gam)$
%is positive semidfinite.
\end{proof}
For the remainder of the work, we assume that 
\eq{\label{eq:blam1}\eta> \nu m=:\bar\eta} 
so that the mapping
$(\beta,\gam)\mapsto \LL_{\emu}(\bg,\tbg)$ %\hLL((\beta,\gam),\tgam)$
is strongly convex with 
%$\nabla_{\bg\bg}^2\hLL(\bg,\tgam)$ 
positive definite Hessian
regardless of the 
choice of $\tgam\in\R^q$. With this in mind, the function $\uemu$ defined by
\eqref{eq:value_function_definition} resembles a Moreau envelope.
%of the convex function $\hLL+\phi_\mu$.
However, this is misleading since, in particular,
we are not even assured of the existence of solutions to
the optimization problem defining $\uemu$.

@@

\begin{theorem}[Convergence of Algorithms \ref{alg:MSR3} and \ref{alg:pgd with bt}]\label{thm:convergence}
Let the assumptions of Theorem \ref{thm:relaxed existence} hold, and
let $\Phi_\emu$ be as defined in \eqref{eq:varphi}. 
%and suppose
%$\gammax\in\R^q_{++}$. 
Let $\{\tbgk\}$ be a sequence generated
by either Algorithm \ref{alg:MSR3} or \ref{alg:pgd with bt} 
with parameters 
$\theta\in(0,1),\ \tau\in(0,1)$,
$\eta>0,\ \mu>0$, $\eps_{\mbox{\tiny Tol}}=0$,
$t_0>0$, and $\gammax>\tgam^0$. 
Then, given $\zeta>\uemu(\tbeta^0,\tgam^0)$ there is an 
$L_{(\emu,\gammax,\zeta)}>0$ such that $\nabla\uemu$ is 
$L_{(\emu,\gammax,\zeta)}$-smooth over 
$\widetilde\VV((\emu,\gammax,\zeta,1)$. 
In Algorithm \ref{alg:pgd with bt}, replace 
$L_\emu$ with $L_{(\emu,\gammax,\zeta)}$ and set
 \[
 \begin{aligned}
 M&:=\begin{cases}
 \alf(1-\alf\frac{L_{(\emu,\gammax,\zeta)}}{2}),&
 \mbox{in Algorithm \ref{alg:pgd with bt}},
 \\
 \frac{2t_0\theta^2(1-\tau)}{\max\{2\theta(1-\tau),t_0L_{(\emu,\gammax,\zeta)}\}},&
 \mbox{in Algorithm \ref{alg:MSR3}},
 \end{cases}
 \qquad\text{and}
 \\
 r&:=
 \begin{cases}
 \alf,&
 \mbox{in Algorithm \ref{alg:pgd with bt}},
 \\
t_0,&
 \mbox{in Algorithm \ref{alg:MSR3}}.
 \end{cases}
 \end{aligned}\]
Then either $\tgam^k>\gammax$ after a finite number of iterations and the algorithms terminate, or the 
following hold:
\begin{enumerate}
\item
The sequence $\Phi_\emu\tbgk$ is nondecreasing.
In addition, \[\Phi_\emu(\tbgk)\tbeta^{k+1},\tgam^{k+1})<\Phi_\emu\tbgk\]
if and only if $\tbgk$ is not a stationary point of \eqref{eq:relax3}.
\item
$\norm{\tbgk-\prox_{r\tR}(\tbgk-r\nabla u_\emu\tbgk)}\rightarrow 0$ with
\[
\min_{i=0,1,\dots,k}\norm{\tbgk-\prox_{r\tR}(\tbgk-r\nabla u_\emu\tbgk)}
\le \frac{\sqrt{\Phi_\emu(\tbeta^0,\tgam^0)-\Phi_\emu^\mathsf{OPT}}}
{\sqrt{M(k+1)}},
\]
where $\Phi_\emu^\mathsf{OPT}:=\inf \Phi_\emu$.
\item
All limit points of the sequence $\{\tbgk\}$ are stationary points of
problem \eqref{eq:relax3}.
\end{enumerate}
\end{theorem}
\begin{proof}
As observed prior to the statement of the theorem, 
both Algorithm \ref{alg:MSR3} and \ref{alg:pgd with bt}
behave as if they were applied to the the functions 
$\hat u_\emu$ and $\hR$ defined in \eqref{eq:hats}. 
It was also shown that the functions $\hat u_\emu$ and $\hR$
satisfy the Assumptions (A)-(C) required by \cite[Theorem 10.15]{AB17}.
Hence, the consequences of \cite[Theorem 10.15]{AB17} hold.
By translating the notion of \cite[Theorem 10.15]{AB17} to that of this work,
we obtain the result.
\end{proof}

@@

\noindent
which is motivated by the expression for $\nabla^2\LL_{\emu}\bg$ given in
\eqref{eq:hess LL}. That is, we simply drop the negative semi-definite term
$-\sum_{i=1}^m\begin{bmatrix}
0&0\\ 0& \half(Z_i^T\Omega_i(\gam)^{-1}Z_i)^{\circ2}
\end{bmatrix}$.
With this modification, the subproblems we solve are strongly convex for all $\eta>0$.
Consequently, the problem of initializing $\eta$ is less problematic. Our numerical experiments indicate that the performance of the algorithm is robust with respect to 
$\eta$. For this reason, we choose an initial value for $\eta$ and 
then leave it fixed over all iterations. Our method for choosing $\eta$ is described
in \cite[Section 4, Figure 5]{sholokhov2022relaxation}. Briefly, we maximize the 
Baysian Information Criterion (BIC) over a grid of values for $\eta$. 
 The resulting BIC response curve shows 
that the method is robust with respect to the 
 choice of $\eta$ and choosing $\eta\in [1,10]$ yields 
 accurate solutions for our selected test problems. 
 Once $\eta$ is fixed the
 PGD algorithm can be applied to solve the problem \eqref{eq:relax3}
 for decreasing values of $\mu$.
